---
title: "Full Figure 2 with cowplot"
output: 
     html_notebook:
          toc: true
          code_folding: hide
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(readr)
library(dplyr)
library(scales)
library(readxl)
library(cowplot)
```


# A: Age
```{r}
library(readxl)
```


```{r}
Early_JP <- read_excel("U:/Box Sync/ProjectDocs/2017_DistantCRC/JoinPoint/Rate_StageEarly.xlsx") %>% rename("Year" = "X Value", "Model" = "Modeleded Y Value", "Rate" = "Observed Y Value") %>% mutate(Year = as.integer(Year), Model = as.numeric(Model), Rate = as.numeric(Rate), AgeCategory = "Early")
Middle_JP <- read_excel("U:/Box Sync/ProjectDocs/2017_DistantCRC/JoinPoint/Rate_StageMiddle.xlsx") %>% rename("Year" = "X Value", "Model" = "Modeleded Y Value", "Rate" = "Observed Y Value") %>% mutate(Year = as.integer(Year), Model = as.numeric(Model), Rate = as.numeric(Rate), AgeCategory = "Middle")
Late_JP <- read_excel("U:/Box Sync/ProjectDocs/2017_DistantCRC/JoinPoint/Rate_StageLate.xlsx") %>% rename("Year" = "X Value", "Model" = "Modeleded Y Value", "Rate" = "Observed Y Value") %>% mutate(Year = as.integer(Year), Model = as.numeric(Model), Rate = as.numeric(Rate), AgeCategory = "Late")

AllAges_JP <- rbind(Early_JP, Middle_JP, Late_JP)
AllAges_JP$AgeCategory <- factor(AllAges_JP$AgeCategory, levels = c("Early","Middle","Late"))
```


# Plots

```{r}
colorclasses <- c("Localized" = "#016b0b","Distant" = "#be0306","Regional" = "#ffb401")

shapes <- c(8, 15, 17)
lwds <- c(3, 4, 2, 1)

```


## Rate of disease stage within Age Category
```{r}

Figure2A <- ggplot() + 
     geom_point(data=AllAges_JP, aes(x=Year, y=Rate, col=Stage, group=Stage, shape=Stage), size=2) + 
     geom_line(data = AllAges_JP, aes(x=Year, y=Model, group=Stage, col=Stage)) + 
     facet_wrap( ~ AgeCategory, scales = "free_y") + 
     scale_x_continuous(breaks=c(1998, 2000, 2005, 2010, 2014), labels = c("1998", "2000", "2005", "2010", "2014"), minor_breaks = seq(1998,2014,1)) + 
     ylab("Incidence rate (per 100,000)")
 
     
Figure2A_themed <- Figure2A + 
     scale_color_manual(name="Stage",values = colorclasses) +
     scale_shape_manual(values=shapes) + 
     theme_classic() + 
     theme(
     legend.position = "none",
     axis.title = element_text(size = 12, face="bold"),
     axis.text = element_text(size=10),
     axis.text.x = element_text(angle=45, vjust=0.5)
     )

Figure2A_themed
```

# B: Sidedness
```{r}
Prox_JP <- read_excel("U:/Box Sync/ProjectDocs/2017_DistantCRC/JoinPoint/Rate_StageProximal.xlsx") %>% rename("Year" = "X Value", "Rate" = "Observed Y Value", "Model" = "Modeleded Y Value") %>% mutate(Year = as.integer(Year), Rate = as.numeric(Rate), Model = as.numeric(Model), Side = "Proximal")
Dist_JP <- read_excel("U:/Box Sync/ProjectDocs/2017_DistantCRC/JoinPoint/Rate_StageDistal.xlsx") %>% rename("Year" = "X Value", "Rate" = "Observed Y Value", "Model" = "Modeleded Y Value") %>% mutate(Year = as.integer(Year), Rate = as.numeric(Rate), Model = as.numeric(Model), Side = "Distal")

AllSides_JP <- rbind(Prox_JP, Dist_JP)
AllSides_JP$Side <- factor(AllSides_JP$Side, levels=c("Proximal","Distal"))
AllSides_JP$Stage <- factor(AllSides_JP$Stage, levels = c("Localized","Regional","Distant"))
```


#Plots

```{r}
colorclasses <- c("Localized" = "#016b0b", "Distant" = "#be0306", "Regional" = "#ffb401")

shapes <- c(15, 17, 8)
lwds <- c(3, 4, 2, 1)
```


##Rate of Disease Stage by Site Category over time
```{r}
Figure2B <- ggplot(AllSides_JP) + 
     geom_point(aes(x=Year, y=Rate, col=Stage, group=Stage, shape=Stage), size=2) + 
     geom_line(aes(x=Year, y=Model, col=Stage, group=Stage)) + 
     facet_wrap(~ Side) + 
     scale_x_continuous(breaks=c(1998, 2000, 2005, 2010, 2014), labels = c("1998", "2000", "2005", "2010", "2014"), minor_breaks = seq(1998,2014,1)) + 
     ylab("Incidence rate (per 100,000)")

Figure2B_themed <- Figure2B + 
     scale_color_manual(name="Stage",values = colorclasses) +
     scale_shape_manual(values=shapes) + 
     theme_classic() + 
     theme(
     legend.position = "none",
     axis.title = element_text(size = 12, face="bold"),
     axis.text = element_text(size=10),
     axis.text.x = element_text(angle=45, vjust=0.5)
     )

Figure2B_themed

#ggsave(plot=Figure2B_themed, filename="Figure2B.tiff", path="C:/Users/gaugustus/Documents/Success_Docs/Pubs/DistantCRC_BriefComm", device = "tiff", width=16, height=10, units="cm", dpi = 300)
```

# C: Race
```{r, message=FALSE}
CRCSites <- readr::read_delim("K:/SEER_DataTables/SEER_2000-2014_CRC_StageDistribution_Race.txt", "\t", escape_double = FALSE, trim_ws = TRUE)
colnames(CRCSites) <- c("Year","Race", "Stage","Rate", "Count", "Population")

CRCSites2 <- CRCSites %>% filter(!grepl("-", Year)) %>% filter(Stage %in% c("Distant","Localized","Regional","Unknown/unstaged")) %>% filter(Year >= 1998) %>% filter(! Race %in% c("Unknown"))
CRCSites2$Year <- as.integer(CRCSites2$Year)
```



Group by Year, Race, Stage and Age, then summarize by each combination.  Get total cases for each year/race/age group.
```{r}
CRCSites2$Stage <- factor(CRCSites2$Stage, levels=c("Unknown/unstaged","Localized","Regional","Distant"))
CRCSites2$Race <- factor(CRCSites2$Race, levels=c("Black", "American Indian/Alaska Native","Asian or Pacific Islander","White"), labels = c("Black or African American", "American Indian/Alaska Native","Asian or Pacific Islander","White"))
```

# The Data we're working with
```{r}
CRCSites2
```

# Read in joinpoint models
```{r}
library(readxl)
```

```{r}
AI_JPs <- read_excel("U:/Box Sync/ProjectDocs/2017_DistantCRC/JoinPoint/Rate_StageAmericanIndian.xlsx", sheet=3) %>% 
     mutate(Race = "American Indian/Alaska Native") %>% 
     rename("Year" = "X Value", "Model" = "Modeleded Y Value") %>% 
     mutate(Year = as.integer(Year), Model = as.numeric(Model), APC=NULL)
A_JPs <- read_excel("U:/Box Sync/ProjectDocs/2017_DistantCRC/JoinPoint/Rate_StageAsian.xlsx", sheet=3) %>% 
     mutate(Race = "Asian or Pacific Islander") %>% 
     rename("Year" = "X Value", "Model" = "Modeleded Y Value") %>% 
     mutate(Year = as.integer(Year), Model = as.numeric(Model))
B_JPs <- read_excel("U:/Box Sync/ProjectDocs/2017_DistantCRC/JoinPoint/Rate_StageBlack.xlsx", sheet=1) %>% 
     mutate(Race = "Black or African American") %>% 
     rename("Year" = "X Value", "Model" = "Modeleded Y Value") %>% 
     mutate(Year = as.integer(Year), Model = as.numeric(Model))
W_JPs <- read_excel("U:/Box Sync/ProjectDocs/2017_DistantCRC/JoinPoint/Rate_StageWhite.xlsx", sheet=1) %>% 
     mutate(Race = "White") %>% rename("Year" = "X Value", "Model" = "Modeleded Y Value") %>% 
     mutate(Year = as.integer(Year), Model = as.numeric(Model), X__1=NULL, X__2=NULL)

AllRace_JPs <- rbind(AI_JPs, A_JPs, B_JPs, W_JPs)
```



##Rate of stage category by year

```{r}
colorclasses <- c("Localized" = "#016b0b","Distant" = "#be0306","Regional" = "#ffb401")

shapes <- c(15,  17, 8)
lwds <- c(3, 2, 1)

```

```{r}
Races <- list(
     'Black or African American'="Black or\nAfrican American",
     'American Indian/Alaska Native'="Native American\nAlaska Native",
     'Asian or Pacific Islander'="Asian or\nPacific Islander",
     'White' = "White"
)
Races_labeller <- function(variable,value){
  return(Races[value])
}

```

```{r}
Figure2C <- ggplot() + 
     geom_point(data=CRCSites2 %>% filter(Stage != "Unknown/unstaged"), aes(x=Year, y=Rate, col=Stage, group=Stage, shape=Stage), size=2) + 
     geom_line(data=AllRace_JPs, aes(x=Year, y=Model, group=Stage, col=Stage)) +
    facet_grid(. ~ Race, labeller = Races_labeller) +
     scale_x_continuous(breaks=c(1998, 2000, 2005, 2010, 2014), labels = c("1998", "2000", "2005", "2010", "2014"), minor_breaks = seq(1998,2014,1)) + 
     scale_y_continuous() + 
     ylab("Incidence rate (per 100,000)") + 
     ggtitle(" ")

Figure2C_themed <- Figure2C + 
     scale_color_manual(name="Stage", values = colorclasses) +
     scale_shape_manual(values=shapes) + 
     theme_classic() + 
     theme(
     legend.position = "none",
     axis.title = element_text(size = 12, face="bold"),
     axis.text = element_text(size=10),
     axis.text.x = element_text(angle=45, vjust=0.5)
     )

Figure2C_themed

#ggsave(plot=Figure2C_themed, filename="Figure2C.tiff", path="C:/Users/gaugustus/Documents/Success_Docs/Pubs/DistantCRC_BriefComm/", device = "tiff", width=16, height=10, units="cm", dpi = 300)

```

# D: Sex
```{r}
RateFemale_JP <- read_excel("U:/Box Sync/ProjectDocs/2017_DistantCRC/JoinPoint/Rate_StageFemale.xlsx") %>% 
     rename("Year" = "X Value", "Rate" = "Observed Y Value", "Model" = "Modeleded Y Value") %>% 
     mutate(Year = as.integer(Year), Rate = as.numeric(Rate), Model = as.numeric(Model), Sex = "Female")
RateMale_JP <- read_excel("U:/Box Sync/ProjectDocs/2017_DistantCRC/JoinPoint/Rate_StageMale.xlsx") %>% 
     rename("Year" = "X Value", "Rate" = "Observed Y Value", "Model" = "Modeleded Y Value") %>% 
     mutate(Year = as.integer(Year), Rate = as.numeric(Rate), Model = as.numeric(Model), Sex = "Male")

AllSex_JP <- rbind(RateFemale_JP, RateMale_JP)

```


#Plots

```{r}
colorclasses <- c("Localized" = "#016b0b", "Distant" = "#be0306", "Regional" = "#ffb401")

shapes <- c(8, 15, 17)
lwds <- c(3, 4, 2, 1)
```


##Rate of Disease Stage by Site Category over time
```{r}
Figure2D <- ggplot(AllSex_JP) + 
     geom_point(aes(x=Year, y=Rate, col=Stage, group=Stage, shape=Stage), size=2) + 
     geom_line(aes(x=Year, y=Model, col=Stage, group=Stage)) + 
     facet_wrap(~ Sex) + 
     scale_x_continuous(breaks=c(1998, 2000, 2005, 2010, 2014), labels = c("1998", "2000", "2005", "2010", "2014"), minor_breaks = seq(1998,2014,1)) + 
     ylab("Incidence rate (per 100,000)")

Figure2D_themed <- Figure2D + 
     scale_color_manual(name="Stage",values = colorclasses) +
     scale_shape_manual(values=shapes) + 
     theme_classic() + 
     theme(
     legend.position = "none",
     axis.title = element_text(size = 12, face="bold"),
     axis.text = element_text(size=10),
     axis.text.x = element_text(angle=45, vjust=0.5)
     )

Figure2D_themed
```

# Cowplot, put it all together
```{r}
Legend <- get_legend(Figure2B_themed + theme(legend.position = "bottom", 
                                          legend.title = element_text(size=12, face="bold"),
                                          legend.text = element_text(size=10))) 
```



```{r}
prow <- plot_grid(Figure2A_themed, Figure2B_themed, Figure2C_themed, Figure2D_themed, labels = c("A", "B", "C", "D"), hjust = -1, nrow = 2, align = "vh", label_size = 26)
p <- plot_grid(prow, Legend, ncol=1, rel_heights = c(1, .1))
p
#ggsave(plot=p, filename="Figure2.tiff", path="C:/Users/gaugustus/Documents/Success_Docs/Pubs/DistantCRC_BriefComm", device = "tiff", width=33, height=20, units="cm", dpi = 300)

#ggsave(plot=p, filename="Figure2_Proof.tiff", path="C:/Users/gaugustus/Documents/Success_Docs/Pubs/DistantCRC_BriefComm", device = "tiff", width=33, height=20, units="cm", dpi = 72)

ggsave(plot=p, filename="Figure3_1200px.tiff", path="U:/Box Sync/Distant CRC - JNCI Brief Comm", device = "tiff", width=12, height=7.5, units = "in", dpi=100)

ggsave(plot=p, filename="Figure3.tiff", path="U:/Box Sync/Distant CRC - JNCI Brief Comm", device = "tiff", width=12, height=7.5, units = "in", dpi=300)
```



```{r}
sessionInfo()
```

