---
title: "Alter tables for Joinpoint Modeling"
output: html_notebook
---


# Generic function to fix common problems in all datasets.
```{r}
prepTable <- function(file, override = FALSE, outputfilename = NULL){
     library(readr)
     library(dplyr)
     tempfile <- readr::read_tsv(file = file, na = " ")
     newfile <- tempfile %>% filter(`Year of diagnosis` != "2000-2014" & `Summary stage 2000 (1998+)` %in% c("Localized","Regional","Distant") & `Age recode with <1 year olds` != "Unknown")
     newfile$`Summary stage 2000 (1998+)` <- factor(newfile$`Summary stage 2000 (1998+)`, levels = c("Localized","Regional","Distant"), labels = c(1, 2, 3))
     Ages <- newfile %>% select(`Age recode with <1 year olds`) %>% distinct(`Age recode with <1 year olds`) %>% pull(`Age recode with <1 year olds`)
     newfile$`Age recode with <1 year olds` <- factor(newfile$`Age recode with <1 year olds`, levels = Ages, labels = c(0,1:(length(Ages)-1)))
     
     newfile <- newfile %>% arrange(`Summary stage 2000 (1998+)`, `Year of diagnosis`, `Age recode with <1 year olds`) %>% rename("Stage" = "Summary stage 2000 (1998+)", "Year of Diagnosis" = "Year of diagnosis", "Age Group" = "Age recode with <1 year olds")
     if(override == TRUE){
          if(is.null(outputfilename)){
               newfilename <- file
          } else {
               newfilename <- outputfilename
          }
          write_tsv(newfile, na = " ", path = newfilename)
     }
     return(newfile)
}
```

## Stage Only
```{r}
filename <- "K:/SEER_DataTables/wStandardPops/SEER_2000-2014_StageDistribution_StageAge.txt"
outputto <- "U:/ProjectDocs/2017_DistantCRC/ProcessedTables/Rate_StageAge_forJoinpoint.txt"
prepTable(filename, override = TRUE, outputfilename = outputto)
```

## Stage + Sex
```{r}
filename <- "K:/SEER_DataTables/wStandardPops/SEER_2000-2014_StageDistribution_StageAgeSex.txt"
outputto <- "U:/ProjectDocs/2017_DistantCRC/ProcessedTables/Rate_StageAgeSex_forJoinpoint.txt"
prepTable(filename, override = TRUE, outputfilename = outputto)
```

## Stage + Race
```{r}
filename <- "K:/SEER_DataTables/wStandardPops/SEER_2000-2014_StageDistribution_StageAgeRace.txt"
outputto <- "U:/ProjectDocs/2017_DistantCRC/ProcessedTables/Rate_StageAgeRace_forJoinpoint.txt"
prepTable(filename, override = TRUE, outputfilename = outputto)
```

## Stage + Site
```{r}
filename <- "K:/SEER_DataTables/wStandardPops/SEER_2000-2014_StageDistribution_StageAgeSidedness.txt"
outputto <- "U:/Box Sync/ProjectDocs/2017_DistantCRC/ProcessedTables/Rate_StageAgeSite_forJoinpoint.txt"
prepTable(filename, override = TRUE, outputfilename = outputto)
```

# Modify datasets as needed

## Stage Only   
No changes needed

## Stage + Age  
Need to add a column for age category.

### Read in data   
```{r}
filename <- "U:/ProjectDocs/2017_DistantCRC/ProcessedTables/Rate_StageAge_forJoinpoint.txt"
StageAgecat <- read_tsv(filename)
```
### Create Age Category column   
```{r}
StageAgecat$`Age Category` <- NA
for(i in 1:nrow(StageAgecat)){
     # if  c("00 years", "01-04 years", "05-09 years", "10-14 years", "15-19 years", "20-24 years", "25-29 years", "30-34 years", "35-39 years", "40-44 years", "45-49 years")
     if(StageAgecat$`Age Group`[i] %in% c(0,1:10)){
          StageAgecat$`Age Category`[i] <- "Early"
     }
     # else if c("50-54 years", "55-59 years","60-64 years")
     else if(StageAgecat$`Age Group`[i] %in% c(11:13) ){
          StageAgecat$`Age Category`[i] <- "Middle"
     }
     else if(StageAgecat$`Age Group`[i] %in% c(14:18)){
          StageAgecat$`Age Category`[i] <- "Late"
     }
}
table(StageAgecat$`Age Category`)

StageAgecat <- StageAgecat %>% select(Stage, `Age Category`, `Year of Diagnosis`, `Age Group`, `Age-Adjusted Rate`, Count, Population, `Std Pop`)

StageAgecat$`Age Category` <- factor(StageAgecat$`Age Category`, levels = c("Early","Middle","Late"), labels = c(1,2,3)) 

StageAgecat
```

### Combine Age Categories
```{r}
# StageAgecat <- StageAgecat %>% 
#      group_by(`Stage`, `Age Category`, `Year of Diagnosis`) %>% 
#      summarize(`Age-Adjusted Rate` = sum(`Age-Adjusted Rate`),
#                Count = sum(Count),
#                Population = Population[1],
#                `Std Pop` = `Std Pop`[1])
# StageAgecat
```



```{r}
#write_tsv(StageAgecat, "U:/ProjectDocs/2017_DistantCRC/ProcessedTables/Rate_StageAgeAgecat_forJoinpoint.txt")
```

## Stage + Sex   
### Read in data   
```{r}
filename <- "U:/ProjectDocs/2017_DistantCRC/ProcessedTables/Rate_StageAgeSex_forJoinpoint.txt"
StageSex <- read_tsv(filename)
```
### Filter for only Male and Female   

```{r}
StageSex <- StageSex %>% filter(Sex %in% c("Male","Female"))
head(StageSex)
```
### Write to file  
```{r}
#write_tsv(StageSex, filename)
```


## Stage + Race   
### Read in data   
```{r}
filename <- "U:/ProjectDocs/2017_DistantCRC/ProcessedTables/Rate_StageAgeRace_forJoinpoint.txt"
StageRace <- read_tsv(filename)
```
### Change column name and remove unknowns   
```{r}
StageRace <- StageRace %>% filter(`Race recode (W, B, AI, API)` != "Unknown") %>% rename("Race recode" = `Race recode (W, B, AI, API)`)
head(StageRace)
```
### Write over original file   
```{r}
#write_tsv(StageRace, filename)
```

## Stage + Site   
No changes needed.



